---
title: "DAIE CA4 SQLite Database Querying File"
author: "Shane Cunningham"
date: "January, 2023"
output:
  html_document:
    df_print: paged
---



```{r setup, include=FALSE}
## Clear console on each execute and set chunk default to not show code
cat('\014')
#knitr::opts_chunk$set(include = FALSE)
```




```{r include=FALSE}
## Download and load libraries
 if(!require("RSQLite"))
   install.packages("RSQLite")

 if(!require("DBI"))
   install.packages("DBI")

 if(!require("dplyr"))
   install.packages("dplyr")

library(RSQLite)
library(DBI)
library(dplyr)
```

```{r include=FALSE}
## Establish a connection to the MySQL DB

conn = dbConnect(RSQLite::SQLite(),
                            dbname= "daie_ca4_data.sqlite")
```

# 1. SELECT with WHERE, LIKE, and OR

```{r}
## Select from the team member table and search for strings that contain the letters in the sequence of ##'ni' or the role of Lead Tester
dbGetQuery(conn, "SELECT * FROM Team_Member 
                  WHERE Role
                  LIKE '%ni%'
                  OR Role = 'Lead Tester' " )

```

# 2. SELECT with DISTINCT and ORDER BY

```{r}

## select column work_item_Id from Team_work_item and order by descending order based on Team_Id
dbGetQuery(conn, "SELECT DISTINCT work_item_Id
                  FROM Team_work_item 
                  ORDER BY Team_Id DESC
           ")

```

# 3. Inner Join

```{r}
# Inner Join Team and Library with project_Id
dbGetQuery(conn, "SELECT Team.Team_type, Library.Id
                  FROM Team
                  INNER JOIN Library
                  ON Team.project_id = Library.project_id
           ")

# Inner Join Team_Member.Role and asset.Type with Library_Id
dbGetQuery(conn, "SELECT Team_Member.Role, asset.Type
                  FROM Team_Member
                  INNER JOIN asset
                  ON Team_Member.Library_Id = asset.Library_Id
           ")

```

# 4. Subquery with SELECT

```{r}

#Subquery with select from Team_Member table and show team members first and sir names with a Team_Id equal to 2001  

dbGetQuery(conn, "SELECT Team_Id, First_Name, Sir_Name FROM Team_Member
                  WHERE Team_Id IN (SELECT Team_Id
                    FROM Team_Member
                    WHERE Team_Id = 2001
                  )")

```

# 5. SELECT across a date range

```{r}
# Select the projects delivery date between two date/times
dbGetQuery(conn, "SELECT * FROM project
                  WHERE Delivery_date BETWEEN '2025-07-22 00:00:00' AND '2029-07-22 00:00:00' 
           ")

```

```{r include=FALSE}

dbDisconnect(conn) 

```

# Bibliography

admin (2019) UPDATE Query in SQLite, W3schools. Available at: https://www.w3schools.blog/update-query-sqlite (Accessed: 18 January 2023).

Bhalla, D. (no date) ‘dplyr Tutorial : Data Manipulation (50 Examples)’, ListenData. Available at: https://www.listendata.com/2016/08/dplyr-tutorial.html (Accessed: 10 December 2022).
Datatypes In SQLite (no date). Available at: https://www.sqlite.org/datatype3.html (Accessed: 18 January 2023).

Execute a data manipulation statement on a given database connection — dbSendStatement (no date). Available at: https://dbi.r-dbi.org/reference/dbsendstatement (Accessed: 18 January 2023).

How to use DATETIME value in SQLite? (2018) TablePlus. Available at: https://tableplus.com/blog/2018/07/sqlite-how-to-use-datetime-value.html (Accessed: 18 January 2023).

[R beginners]: Use  RSQLite, a fast and easy database in R (2020). Available at: https://www.youtube.com/watch?v=GYfmN2IeY1w (Accessed: 16 January 2023).

R Markdown for a Data Analysis Report (2016). Available at: https://www.youtube.com/watch?v=0bxa7rkur9o (Accessed: 7 December 2022).

R programming for beginners – statistic with R (t-test and linear regression) and dplyr and ggplot (2017). Available at: https://www.youtube.com/watch?v=ANMuuq502rE (Accessed: 20 December 2022).

R SQLite Database Tutorial (no date). Available at: https://www.datacamp.com/tutorial/sqlite-in-r (Accessed: 18 January 2023).

Riederer, Y.X., Christophe Dervieux, Emily (no date) How to read this book | R Markdown Cookbook. Available at: https://bookdown.org/yihui/rmarkdown-cookbook/how-to-read-this-book.html (Accessed: 10 December 2022).

‘rmarkdown-cheatsheet’ (no date).
Solutions - SQLite (NA). Available at: https://solutions.posit.co/connections/db/databases/sqlite/ (Accessed: 18 January 2023).

SQL BETWEEN Operator (no date). Available at: https://www.w3schools.com/sql/sql_between.asp (Accessed: 19 January 2023).

SQL INNER JOIN Keyword (no date). Available at: https://www.w3schools.com/sql/sql_join_inner.asp (Accessed: 19 January 2023).

SQL ORDER BY Keyword (no date). Available at: https://www.w3schools.com/sql/sql_orderby.asp (Accessed: 19 January 2023).

SQL SELECT DISTINCT Statement (no date). Available at: https://www.w3schools.com/sql/sql_distinct.asp (Accessed: 19 January 2023).

SQL Subquery – How to Sub Query in SELECT Statement (2022) freeCodeCamp.org. Available at: https://www.freecodecamp.org/news/sql-subquery-how-to-sub-query-in-select-statement/ (Accessed: 19 January 2023).

SQLite Home Page (no date). Available at: https://www.sqlite.org/index.html (Accessed: 18 January 2023).





